State of the Practice for Lattice Boltzmann Method Software

### 2022-03-18 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Addition of figure of LBM for aorta artery flow

We analyze the development of Computational Fluid Dynamics (CFD) software
packages that use the Lattice Boltzmann Method (LBM). LBM packages form a family
of algorithms for simulating single-phase and multiphase fluid flows, often
incorporating additional physical complexities \citep{chen1998lattice}, such as
reflective and non-reflective boundaries. LBM considers the behaviour of a
collection of particles as a single unit at the mesoscopic scale, which lies
between the nanoscopic and microscopic scales. LBM solvers predict the
positional probability of a collection of particles moving through a lattice
structure following a two step process: i) streaming, where the particles move
along the lattice via links; and, ii) colliding, where energy and momentum is
transferred among particles that collide \citep{bao2011lattice}. As an example
of the output of LBM, Figure~\ref{circularflow} presents the streamlines of
converged flow past a stationary circular cylinder with varying Reynolds
numbers. \wss{Zahra can likely provide a better example.}  LBM has several
advantages over conventional CFD methods, including a simple calculation
procedure, improved parallelization, and robust handling of complex geometries
\citep{ganji2015application}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{./figures/circularflow}
		\caption{Streamlines of flow past a stationary circular cylinder at
		Reynolds number = 10, 20, and 40 \citep{chen2021phase}}
		\label{circularflow}
	\end{center}
\end{figure}

@@

We investigated the research questions by applying the general methodology
summarized in \citet{SmithEtAl2021}.  The specific application of the
methodology to LBM is reviewed in Section~\ref{methodology}, with the full
details in \citep{Michalski2021}.  The current methodology updates the approach
used in prior assessments of domains like Geographic Information Systems
\citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Oceanographic Software \citep{smith2015state}, Seismology software
\citep{SmithEtAl2018}, statistical software for psychology
\citep{SmithEtAl2018_StatSoft} and medical image analysis software
\citep{Dong2021}.  \wss{add citation to medical imaging software, if it is an
option}

>>>>

We analyze the development of Computational Fluid Dynamics (CFD) software
packages that use the Lattice Boltzmann Method (LBM). LBM packages form a family
of algorithms for simulating single-phase and multiphase fluid flows, often
incorporating additional physical complexities \citep{chen1998lattice}, such as
reflective and non-reflective boundaries. LBM considers the behaviour of a
collection of particles as a single unit at the mesoscopic scale, which lies
between the nanoscopic and microscopic scales. LBM solvers predict the
positional probability of a collection of particles moving through a lattice
structure following a two step process: i) streaming, where the particles move
along the lattice via links; and, ii) colliding, where energy and momentum is
transferred among particles that collide \citep{bao2011lattice}. As an example
of an important application of LBM, Figure~\ref{Fig_coarctation} shows the flow
in a defective aorta pre and post medical intervention. LBM has several
advantages over conventional CFD methods, including a simple calculation
procedure, improved parallelization, and robust handling of complex geometries
\citep{ganji2015application}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{./figures/FlowModellingCoarctation.pdf}
		\caption{LBM (and Lattice Poisson-Boltzmann) computed flow (velocity
		magnitude, turbulent kinetic energy and time averaged wall shear stress)
		in an aorta with a coarctation (narrowing of the artery) and a major
		aneurysm (a ballooning of the artery wall) downstream of the
		coarctation, pre and post intravascular stent intervention
		\citep{SadeghiEtAl2020}}.
		\label{Fig_coarctation}
	\end{center}
\end{figure}

@@

We investigated the research questions by applying the general methodology
summarized in \citet{SmithEtAl2021}.  The specific application of the
methodology to LBM is reviewed in Section~\ref{methodology}, with the full
details in \citep{Michalski2021}.  The current methodology updates the approach
used in prior assessments of domains like Geographic Information Systems
\citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Oceanographic Software \citep{smith2015state}, Seismology software
\citep{SmithEtAl2018}, statistical software for psychology
\citep{SmithEtAl2018_StatSoft} and medical image analysis software
\citep{Dong2021}.
